---
title: Shiny Apps for Analyzing Dallas Police Data
author: Sean Conroy
date: '2021-01-14'
slug: shiny-apps-for-analyzing-dallas-police-data
categories:
  - Dallas-Police-Data
tags:
  - shiny
---


To explore the Dallas Crime data more yourself, check out [my Dallas Crime Explorer Shiny app here](https://sean-conroy.shinyapps.io/shinycrimeexplorer/){target="_blank"}.

To explore the Dallas Murder data more yourself, check out [my Dallas Murder Shiny app here](https://sean-conroy.shinyapps.io/shinymurdermap/){target="_blank"}.


## Dallas Trafficking

To explore the Dallas Trafficking data yourself, check out [my Dallas Trafficking Shiny app here](https://sean-conroy.shinyapps.io/shinytraffic/){target="_blank"}.

```{r Load Data,echo = FALSE,eval=FALSE}
library(data.table)
PI <- readRDS("C:/Users/sconroy/Documents/DallasPoliceData/PoliceIncidents1-9-21.RDS")

```

To explore any recorded incidents of human trafficking in the Dallas area, I used the following code:

```{r Extract Traffick,echo = TRUE,eval=FALSE}

setDT(PI)
traffic <- PI[grepl("TRAFFICKING",offincident),]
traffic[,NumIncidentsPerYear := .N,by = "servyr"]

traffic[,rowid := 1:.N]
traffic[,Date := as.Date(substr(date1,1,10))]
traffic[,LatLongStart := regexpr("(",geocoded_column,fixed = TRUE)[1] + 1,by = rowid]
traffic[,LatLongEnd := regexpr(")",geocoded_column,fixed = TRUE)[1] - 1,by = rowid]
traffic[,LatLong := substr(geocoded_column,start = LatLongStart,stop = LatLongEnd)]
traffic[,LatLongComma := regexpr(",",LatLong,fixed = TRUE)[1],by = rowid]
traffic[,Latitude := substr(LatLong,start = 1,stop = LatLongComma - 1)]
traffic[,Longitude := substr(LatLong,start = LatLongComma + 1,stop = 1000)]
traffic[,Longitude := as.numeric(Longitude)]
traffic[,Latitude := as.numeric(Latitude)]

# Manually fix two missing geocodes
traffic[is.na(Latitude) & incident_address == "14040 N STEMMONS SERV",Latitude := 32.93769989950343]
traffic[is.na(Longitude) & incident_address == "14040 N STEMMONS SERV",Longitude := -96.90205446873641]

traffic[is.na(Latitude) & incident_address == "7815 L B J FWY",Latitude := 32.925475281010286]
traffic[is.na(Longitude) & incident_address == "7815 L B J FWY",Longitude := -96.77161085979215]

traffic[,Year := as.factor(servyr)]

```


Here's the code to create the Leaflet map within the Shiny app:

```{r Leaflet,echo = TRUE,eval=FALSE}

        leaflet(data = traffic) %>% 
            addProviderTiles(providers$Stamen.TonerLite,options = providerTileOptions(noWrap = TRUE)) %>% 
                addCircleMarkers(lng = ~Longitude, lat = ~Latitude, weight = 2, radius = input$MarkerSize,
                                 fillOpacity = input$MarkerOpacity,fillColor = ~pal(Year),color = "gray",stroke = 1,
                                 popup = ~paste0("DPD Incident Date: ",traffic$Date,", Incident Address: ",incident_address,
                                                 ", Victim Sex: ",traffic$compsex,
                                                ", Victim Age: ",traffic$compage,", MO: ",traffic$mo)) %>%
                addLegend(position = "bottomleft",
                          pal = pal, values = ~traffic$Year,
                          title = "Legend",
                          opacity = 1)

```


The full code behind the Shiny App is [on GitHub here](https://github.com/meystingray/meystingray.github.io/tree/master/ShinyTraffic){target="_blank"}.
