---
title: Mapping Murders in Dallas
author: Sean Conroy
date: '2020-12-22'
slug: mapping-murders-in-dallas
categories:
  - Dallas-Police-Data
tags: []
---

To explore the Dallas Murder data more yourself, check out [my Shiny app here](https://sean-conroy.shinyapps.io/shinymurdermap/){target="_blank"}.

The code behind this blog post is [on GitHub here](https://github.com/meystingray/meystingray.github.io/blob/master/content/post/2020-12-22-mapping-murders-in-dallas/index.RMD){target="_blank"}.

Note, 1-11-2020: Data has been refreshed to include all of 2020.
```{r Load Library, echo = FALSE}
pkg <- c("data.table","ggplot2","stringr","dplyr","gganimate","RSocrata","ggmap")
suppressWarnings(suppressPackageStartupMessages(
    invisible(lapply(pkg, function(x) require(x, character.only = T, quietly = T)))
))

library(gganimate)
library(ggmap)
library(gifski)

#PI <- read.socrata("https://www.dallasopendata.com/resource/qv6i-rri7.csv")
#saveRDS(PI,file = "C:/Users/sconroy/Desktop/Debug/PoliceIncidents12-17-20.RDS")

PI <- readRDS("C:/Users/sconroy/Documents/DallasPoliceData/PoliceIncidents1-9-21.RDS")
setDT(PI)

Murder <- PI[grepl("MURDER",offincident) | grepl("HOMICIDE",offincident) | grepl("MURDER",ucr_offense) | 
                 grepl("HOMICIDE",nibrs_crime_category) | grepl("MURDER",nibrs_crime),]



# Extract Lat / Long frm geocoded column
Murder[,rowid := 1:.N]
Murder[,Date := as.Date(substr(date1,1,10))]
Murder[,MonthDate := as.Date(paste0(format(Date,"%Y-%m"),"-01"))]
Murder[,LatLongStart := regexpr("(",geocoded_column,fixed = TRUE)[1] + 1,by = rowid]
Murder[,LatLongEnd := regexpr(")",geocoded_column,fixed = TRUE)[1] - 1,by = rowid]
Murder[,LatLong := substr(geocoded_column,start = LatLongStart,stop = LatLongEnd)]
Murder[,LatLongComma := regexpr(",",LatLong,fixed = TRUE)[1],by = rowid]
Murder[,Latitude := substr(LatLong,start = 1,stop = LatLongComma - 1)]
Murder[,Longitude := substr(LatLong,start = LatLongComma + 1,stop = 1000)]
Murder[,Longitude := as.numeric(Longitude)]
Murder[,Latitude := as.numeric(Latitude)]
Murder[,Year := as.factor(servyr)]
MurderWithLatLong <- Murder[!is.na(Latitude) & !is.na(Longitude) & !is.na(Date) & !is.na(servyr),
                            .(Date,Longitude,Latitude,Year,servyr)]
Murders2020 <- MurderWithLatLong[Date > as.Date("2020-01-01"),.(Date,Latitude,Longitude)]

setorder(Murder,Date)
# Shootings by Year on Map
#ggplot(Murder) + geom_point(aes(x = Longitude,y = Latitude,color = signal)) + facet_wrap(~ servyr) + ylab(NULL)
#ggplot(Murder[signal == "19 - SHOOTING",]) + geom_point(aes(x = Longitude,y = Latitude)) + facet_wrap(~ servyr) + ylab(NULL)

```

```{r Load GoogleMap, echo = FALSE,message = FALSE}
# Load our API key for Google

apiKey <- fread("C:/Users/sconroy/Documents/DallasPoliceData/APIkey.key")
apiKey <- names(apiKey)
register_google(key = apiKey)

# get Bounding Box for Data Set
lats <- c(min(MurderWithLatLong$Latitude),max(MurderWithLatLong$Latitude))
lons <- c(min(MurderWithLatLong$Longitude),max(MurderWithLatLong$Longitude))
bb <- make_bbox(lon=lons,lat=lats,f=0.05)


# Get map of Dallas from Google API
DallasMapZoom10 <- get_map(location = "Dallas", zoom = 10, maptype = "roadmap",source = "google",messaging = FALSE)
DallasMapZoom11 <- get_map(location = bb, zoom = 11, maptype = "roadmap",source = "google",messaging = FALSE)
#DallasMapZoom12 <- get_map(location = "Dallas", zoom = 12, source = "google")

cda <- get_map(bb,maptype = "roadmap",source = "google")

```


All Dallas Murders since 2014, Color by Year

```{r DallasAll, echo = FALSE,message = FALSE}
ggmap(DallasMapZoom11,extent = "device") + 
    geom_point(data = MurderWithLatLong, aes(x = Longitude, y = Latitude,fill = Year),
               shape = 21, size = 2, stroke = 1,color = "blue") +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    ggtitle("Dallas Murders since 2014") + 
    scale_fill_manual(values = c("black", "dark gray", "blue", "purple", "green","orange","red","yellow"))
```


Let's take a look at the Dallas murders on a map for each year.

```{r Plot Using the Maps, echo = FALSE,message = FALSE}
ggmap(DallasMapZoom11,extent = "normal") + 
    geom_point(data = MurderWithLatLong, aes(x = Longitude, y = Latitude), size = 1,color = "brown") +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    facet_wrap(. ~ servyr) + ggtitle("Dallas Murders by Year")
```


Let's take a look at the murders just for 2020.

```{r 2020Only, echo = FALSE,message = FALSE}

# suppressWarnings(
#     ggmap(DallasMapZoom11,extent = "device") + 
#         geom_point(data = MurderWithLatLong[Date > as.Date("2020-01-01"),], aes(x = Longitude, y = Latitude), 
#                shape = 21, colour = "black", fill = "yellow", size = 3, stroke = 2)
# )

ggmap(cda) + theme_bw() +
    geom_point(data = Murders2020, aes(x = Longitude, y = Latitude), 
               shape = 21, colour = "black", fill = "yellow", size = 3, stroke = 1) + 
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())


```


Cluster analysis: what are the centers of the murder areas?  Let's try K-means with 10 clusters.

```{r Cluster Analysis, echo = FALSE,message = FALSE}

# Try to do Cluster Analysis
KM <- kmeans(Murders2020[!is.na(Latitude) & !is.na(Longitude),.(Latitude,Longitude)],centers = 10)
#KM$centers

# Plot the clusters
suppressWarnings(
    ggmap(DallasMapZoom11,extent = "device") + 
        geom_point(data = Murders2020, aes(x = Longitude, y = Latitude), 
           colour = "black", fill = "black", size = 2) + 
    geom_point(data = as.data.frame(KM$centers), aes(x = Longitude, y = Latitude), 
               shape = 21, colour = "red", fill = "yellow", size = 30, stroke = 2,alpha = 0.3)

)

```


Animate the murders by year.

```{r Plot the anmination, echo = FALSE,message = FALSE}

p <- ggmap(DallasMapZoom11,extent = "device") +
    geom_point(data = MurderWithLatLong, aes(x = Longitude, y = Latitude,group = seq_along(servyr)), 
               size = 3,shape = 21,color = "black",fill = "red",stroke = 1) +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

anim <- p + transition_states(servyr,
                              transition_length = 0,
                              state_length = 1)# + ease_aes()

anim <- anim + ggtitle('Now showing {closest_state}') #,subtitle = 'Frame {frame} of {nframes}')
    
animate(anim,renderer = gifski_renderer(), height = 600, width = 600)

```



Package "ggmap" citation:
D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2. The R Journal, 5(1), 144-161.
URL http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf